Fractional Decomposition Trees: A tool for studying mixed-integer-program integrality gaps

### 2018-11-21 IPCO format/binaryFDT.tex
lemma2 proof
fix

\section{Finding a feasible solution}\label{domTOIP}
Consider a formulation instance $I=(A,G,b)$. Define sets $S(I)$ and $P(I)$ as in (\ref{S'}) and (\ref{P'}), respectively. For simplicity in the notation we denote $P(I),S(I),$ and $g(I)$ with $P$, $S$, and $g$ for this section and the next section. Also, for both sections we assume $t=|\spp(x)|$. Without loss of generality we can assume $x_i = 0$ for $i=t+1,\ldots,n$.

@@

		\begin{equation}\label{splitting}
		x'\geq \sum_{i=1}^{k}\frac{\theta_i}{g} \tilde{x}^i
		={\sum_{i\in [k]: \tilde{x}^i_\ell =0}\frac{\theta_i}{g} \tilde{x}^i}+{\sum_{i\in [k]: \tilde{x}^i_\ell =1}\frac{\theta_i}{g} \tilde{x}^i}
		\end{equation}
		For $j=0,1$, let $(x^j,y^j) = \sum_{i\in [k]:\tilde{x}^i_\ell=j} \frac{\theta_i}{g}(\tilde{x}^i,\tilde{y}^i)$. Also let $\lambda_0=\sum_{i\in [k]: \tilde{x}^i_\ell =0}\frac{\theta_i}{g}$ and $\lambda_1 = \sum_{i\in [k]: \tilde{x}^i_\ell =1}\frac{\theta_i}{g}$. Note that $\lambda_0+\lambda_1 = \frac{1}{g}$. Constraint (\ref{packing}) is satisfied by Inequality (\ref{splitting}). Also, for $j=0,1$ we have
		\begin{equation}
		Ax^j+Gy^j = \sum_{i\in[k], \tilde{x}^i_\ell = j} \frac{\theta_i}{g} (A\tilde{x}^i + G\tilde{y}^i) \geq b \sum_{i\in[k], \tilde{x}^i_\ell = j} \frac{\theta_i}{g} = b\lambda_j.
		\end{equation}
		Hence, Constraints (\ref{feasibility}) holds. Constraint (\ref{branchcoordinate}) also holds since $x^0_\ell$ is obviously $0$ and $x^1_\ell= \sum_{i\in [k]: \tilde{x}^i_\ell = 1}\frac{\theta_i}{g}= \lambda_1$. The rest of the constraints trivially hold. 
	\end{proof}
	This concludes the proof.	
\end{proof}

@@

	Notice that $\theta = \gamma$ is in fact a feasible solution to $\prun(L')$. Let $\theta^*$ be the optimal vertex point solution to this LP. Since the problem is in $\bbbr^{2k}$,  $\theta^*$ has to satisfy $2k$ linearly independent constraints at equality. However, there are only $t$ constraints of type (\ref{packs}). Therefore, there are at most $t$ coordinates of $\theta^*_j$ that are non-zero. We claim that $L_{i+1}$ which consists of $(x^j,y^j)$ for $j=1,\ldots,2k$ and their corresponding multipliers $\theta^*_j$ satisfy the properties in the statement of the lemma. Notice that, we can discard the nodes in $L_{i+1}$ that have $\theta^*_j=0$, so $|L_{i+1}| \leq t$. Also, since $\theta^*$ is optimal and $\gamma$ is feasible for $\prun(L')$, we have $\sum_{j=1}^{k} \theta^*_j \geq \sum_{j=1}^{2k}\gamma_j \geq \frac{1}{g^{i+1}}$. 
\end{proof}
\paragraph{From leaves of FDT to feasible solutions}
The leaves of the FDT tree,  $L_t$, have the property that every solution $(x,y)$ in $L_t$ has $x\in\{0,1\}^n$ and $(x,y)\in \dom(P)$. By applying Lemma \ref{domlemma} we can obtain a point $(x',y')\in S$ such that $x'\leq x$. This conclude the description of the FDT algorithm and proves Theorem \ref{binaryFDT}.

>>>>

\section{Finding a feasible solution}\label{domTOIP}
Consider a formulation instance $I=(A,G,b)$. Define sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. And assume $S(I),P(I)\in [0,1]^n\times \bbbr^p$. For simplicity in the notation we denote $P(I),S(I),$ and $g(I)$ with $P$, $S$, and $g$ for this section and the next section. Also, for both sections we assume $t=|\spp(x)|$. Without loss of generality we can assume $x_i = 0$ for $i=t+1,\ldots,n$.

@@

		\begin{equation}\label{splitting}
		x'\geq \sum_{i=1}^{k}\frac{\theta_i}{g} \tilde{x}^i
		={\sum_{i\in [k]: \tilde{x}^i_\ell =0}\frac{\theta_i}{g} \tilde{x}^i}+{\sum_{i\in [k]: \tilde{x}^i_\ell =1}\frac{\theta_i}{g} \tilde{x}^i}
		\end{equation}
		For $j=0,1$, let $(x^j,y^j) = \sum_{i\in [k]:\tilde{x}^i_\ell=j} \frac{\theta_i}{g}(\tilde{x}^i,\tilde{y}^i)$. Also let $\lambda_0=\sum_{i\in [k]: \tilde{x}^i_\ell =0}\frac{\theta_i}{g}$ and $\lambda_1 = \sum_{i\in [k]: \tilde{x}^i_\ell =1}\frac{\theta_i}{g}$. Note that $\lambda_0+\lambda_1 = \frac{1}{g}$. Constraint (\ref{packing}) is satisfied by Inequality (\ref{splitting}). Also, for $j=0,1$ we have
		\begin{equation}
		Ax^j+Gy^j = \sum_{i\in[k], \tilde{x}^i_\ell = j} \frac{\theta_i}{g} (A\tilde{x}^i + G\tilde{y}^i) \geq b \sum_{i\in[k], \tilde{x}^i_\ell = j} \frac{\theta_i}{g} = b\lambda_j.
		\end{equation}
		Hence, Constraints (\ref{feasibility}) holds. Constraint (\ref{branchcoordinate}) also holds since $x^0_\ell$ is obviously $0$ and $x^1_\ell= \sum_{i\in [k]: \tilde{x}^i_\ell = 1}\frac{\theta_i}{g}= \lambda_1$. The rest of the constraints trivially hold. 
	\end{proof}
	This concludes the proof of Lemma \ref{LPClemma}.	
\end{proof}

@@

	Notice that $\theta = \gamma$ is in fact a feasible solution to $\prun(L')$. Let $\theta^*$ be the optimal vertex point solution to this LP. Since the problem is in $\bbbr^{2k}$,  $\theta^*$ has to satisfy $2k$ linearly independent constraints at equality. However, there are only $t$ constraints of type (\ref{packs}). Therefore, there are at most $t$ coordinates of $\theta^*_j$ that are non-zero. We claim that $L_{i+1}$ which consists of $(x^j,y^j)$ for $j=1,\ldots,2k$ and their corresponding multipliers $\theta^*_j$ satisfy the properties in the statement of the lemma. Notice that, we can discard the nodes in $L_{i+1}$ that have $\theta^*_j=0$, so $|L_{i+1}| \leq t$. Also, since $\theta^*$ is optimal and $\gamma$ is feasible for $\prun(L')$, we have $\sum_{j=1}^{k} \t\label{key}heta^*_j \geq \sum_{j=1}^{2k}\gamma_j \geq \frac{1}{g^{i+1}}$. 
\end{proof}
\paragraph{From leaves of FDT to feasible solutions}
The leaves of the FDT tree,  $L_t$, have the property that every solution $(x,y)$ in $L_t$ has $x\in\{0,1\}^n$ and $(x,y)\in \dom(P)$. By applying Lemma \ref{domlemma} we can obtain a point $(x',y')\in S$ such that $x'\leq x$. This conclude the description of the FDT algorithm and proves Theorem \ref{binaryFDT}.
